---
title: "Peer Assessment1 of Reproducible Research"
output: html_document
---


Loading and Preprocessing the Data

```{r, echo=TRUE}
data <- read.table("activity.csv", header=T, quote="\"", sep=",")

# convert date to date data type
data$date <- as.Date(data$date) 

```

what is the mean total number of steps taken per day?
```{r, echo=TRUE}
data.ignore.na<-na.omit(data)
daily.steps<-rowsum(data.ignore.na$steps,format(data.ignore.na$date,'%Y-%m-%d'))
daily.steps<-data.frame(daily.steps)
names(daily.steps)<-("steps")
```
plot histogram of total number of steps taken each day
```{r, echo=TRUE}
hist(daily.steps$steps,main=" ",breaks=10,xlab="total number of steps taken daily")
```
calculate mean and median of steps
```{r, echo=TRUE}
mean(daily.steps$steps)
median(daily.steps$steps)
```
what is the daily average activity pattern?
```{r, echo=TRUE}
library(plyr)
interval.mean.steps<-ddply(data.ignore.na,~interval,summarise,mean=mean(steps))
```
plot the time series of 5-minute interval and average of number of steps taken, averaged acorss all days.
```{r, echo=TRUE}
library(ggplot2)
qplot(x=interval,y=mean,data=interval.mean.steps,geom="line",xlab="5 minute interval",ylab="number of step count",main="average number of steps taken averaged acorss all days")
```
input missing values
```{r, echo=TRUE}
library(sqldf)
varNA<-sqldf('select d.* from "data" as d where d.steps is null order by d.date,d.interval')
 NROW(varNA)            
```
Implement a strategy to fill the missing data values in the data set.
```{r, echo=TRUE}
fillNA<-sqldf('select d.*,i.mean from "interval.mean.steps" as i join "data" as d on d.interval = i.interval order by d.date,d.interval')
fillNA$steps[is.na(fillNA$steps)]<-fillNA$mean[is.na(fillNA$steps)]
```
Below, plot the histogram and calculate mean and median with missing values filled in.
```{r, echo=TRUE}
fillNA.total.steps<-as.integer(sqldf('select sum(steps) from fillNA'))

fillNA.total.steps.by.date <- sqldf(' 
    SELECT date, sum(steps) as "steps" 
    FROM fillNA GROUP BY date 
    ORDER BY date') 

daily.steps <- sqldf('   
    SELECT date, steps
    FROM "fillNA.total.steps.by.date"
    ORDER BY date')

```
make a histogram of total number of steps taken each day.
```{r, echo=TRUE}
hist(daily.steps$steps,main=" ",breaks=10,xlab="after input NA--total number of steps taken daily")
```
calculate the mean and median total number of steps taken per day.
```{r, echo=TRUE}
mean.steps.per.day<-as.integer(fillNA.total.steps/NROW(fillNA.total.steps.by.date))
mean.steps.per.day

median.steps.per.day<-median(fillNA.total.steps.by.date$steps)
median.steps.per.day
```
are there any differences in activity pattern between weekdays and weekends ?
```{r, echo=TRUE}
fillNA$weektime<-as.factor(ifelse(weekdays(fillNA$date) %in% c("Saturday","Sunday"),"weekend","weekday"))
var2<-sqldf('select interval,avg(steps) as "mean.steps",weektime from fillNA group by weektime,interval order by interval')
```
make a plot containing a time series plot of the 5 minute interval(x-axis) and average number of steps taken, averaged across all weekdays or weekend days(y-axis).
```{r, echo=TRUE}
library(lattice)
plot<-xyplot(mean.steps~interval|factor(weektime),data=var2,type='l',main="average number of steps taken \naveraged across all weekdays or weekend days",xlab="5 minute interval",ylab="average number of steps taken")
print(plot)

```
